\mutagen: Coverage-Guided, Property-Based Testing using Exhaustive Type-Preserving Mutations

### 2021-04-07 papers/oopsla-21/main.tex
Wrote introduction

%%
%% The "title" command has an optional parameter,
%% allowing the author to define a "short title" to be used in page headers.
\title[\mutagen]{\mutagen: Reliable Coverage-Guided, Property-Based Testing using
  Exhaustive Well-Typed Mutations}

@@

Nullam eu ante vel est convallis dignissim. Fusce suscipit, wisi nec facilisis
facilisis, est dui fermentum leo, quis tempor ligula erat quis odio. Nunc porta
vulputate tellus. Nunc rutrum turpis sed pede. Sed bibendum. Aliquam posuere.
Nunc aliquet, augue nec adipiscing interdum, lacus tellus malesuada massa, quis
varius mi purus non odio. Pellentesque condimentum, magna ut suscipit hendrerit,
ipsum augue ornare nulla, non luctus diam neque sit amet urna. Curabitur
vulputate vestibulum lorem. Fusce sagittis, libero non molestie mollis, magna
orci ultrices dolor, at vulputate neque nulla lacinia eros. Sed id ligula quis
est convallis tempor. Curabitur lacinia pulvinar nibh. Nam a sapien.

The contributions of this work are:

\begin{itemize}
\item Nullam eu ante vel est convallis dignissim. Fusce suscipit, wisi nec
  facilisis facilisis, est dui fermentum leo, quis tempor ligula erat quis odio.
  Nunc porta vulputate tellus. Nunc rutrum turpis sed pede. Sed bibendum.
  Aliquam posuere. Nunc aliquet, augue nec adipiscing interdum, lacus tellus
  malesuada massa, quis varius mi purus non odio.
\item Pellentesque condimentum,
  magna ut suscipit hendrerit, ipsum augue ornare nulla, non luctus diam neque
  sit amet urna. Curabitur vulputate vestibulum lorem. Fusce sagittis, libero
  non molestie mollis, magna orci ultrices dolor, at vulputate neque nulla
  lacinia eros. Sed id ligula quis est convallis tempor. Curabitur lacinia
  pulvinar nibh. Nam a sapien.
\item Nullam eu ante vel est convallis dignissim. Fusce suscipit, wisi nec
  facilisis facilisis, est dui fermentum leo, quis tempor ligula erat quis odio.
  Nunc porta vulputate tellus. Nunc rutrum turpis sed pede. Sed bibendum.
  Aliquam posuere. Nunc aliquet, augue nec adipiscing interdum, lacus tellus
  malesuada massa, quis varius mi purus non odio.
\item Pellentesque condimentum, magna ut suscipit hendrerit, ipsum augue ornare
  nulla, non luctus diam neque sit amet urna. Curabitur vulputate vestibulum
  lorem. Fusce sagittis, libero non molestie mollis, magna orci ultrices dolor,
  at vulputate neque nulla lacinia eros. Sed id ligula quis est convallis
  tempor. Curabitur lacinia pulvinar nibh. Nam a sapien.
\end{itemize}

@@

In this section we briefly introduce the concept of Property-based Testing along
with \quickcheck, the most popular tool of this sort and often considered as the
baseline when comparing PBT algoritms.
%
Moreover, we describe the ideas and limitations behind \fuzzchick, which
enhances the PBT approach with execution information and serves as the
foundation for \mutagen.
%
\fuzzchick is implemented as an extension of \quickchick, Coq's own
reimplementation of \quickcheck.

@@

To alleviate the problem of testing properties with sparse preconditions while
using simple automatically derived random generators, \fuzzchick introduces
\emph{coverage-guided property-based testing} (CGPT) by enhancing the testing
process with two key characteristics:
%
\begin{inparaenum}
\item \emph{target code instrumentation}, to capture execution information from
  each test case; and
\item \emph{high-level, well-typed mutations}, to produce syntactically valid
  test cases by altering existing ones at the datatype level.
\end{inparaenum}

@@

The next subsections describe \fuzzchick's testing loop and well-typed mutations
in detail.

@@

\subsubsection{Well-typed mutations}

@@

However, the authors acknowledge that certain parts of their implementation have
room for improvement, especially when it comes to the mutators design.
%
Moreover, when we recreated the evaluation of the IFC stack machine case study
(described in detail in Section \ref{sec:casestudies}), we found that after 30
runs (as opposed to the 5 runs used original by
\citeauthor{lampropoulos2019coverage}), \emph{\fuzzchick was only able to find 5
  out of the 20 injected bugs with a failure rate of 1}, the hardest one being
found only around 13\% of the time after an hour of testing.
%
These results are presented in detail in Section \ref{sec:evaluation}.

At the light of these observations, we identified several aspects of \fuzzchick
that can be improved upon --- and that constitute the main goal of this work.
%
In no particular order:
%
\begin{itemize}
\item \emph{Mutators distribution:} if we inspect the mutator defined in Fig.
  \ref{fig:fuzzchick:mutator}, there are two compromises that the authors of
  \fuzzchick adopted for the sake of simplicity.
  %
  On one hand, deep recursive mutations are very unlikely, since their
  probability decreases multiplicatively with each recursive call.
  %
  For instance, mutating a subexpression that lies on the third level of a
  \texttt{Tree} happens with a probability smaller than $(1/6)^3 = \sim 0.0046$,
  and this only worsens as the type of the mutated value becomes more complex.
  %
  Hence, \fuzzchick mutators can only be effectively used to transform to
  shallow data structures, potentially excluding interesting applications that
  might require producing deeper valid values, e.g., programming languages,
  network protocols interactions, etc.
  %
  Ideally, mutations should be able to happen on every subexpression of the
  input seed in a reasonable basis.

@@

\item \emph{Power schedule:} It is not clear how the power schedule used to
  assign energy to each mutable test case in \fuzzchick works in the context of
  high-level well-typed mutations.
  %
  If it assigns too much energy to certain not-so-interesting seeds, some bugs
  might not be discovered in a timely basis.
  %
  Conversely, assigning too little energy to interesting test cases might cause
  that some bugs cannot be discovered at all unless the right mutation happens
  within the small available energy window --- randomly generating the same test
  case later on does not help, as it becomes uninteresting based on historic
  trace information.

@@

\begin{wrapfigure}{r}{0.5\textwidth}
\vspace{-10pt}
\footnotesize
\begin{tabular}{ccccc}
\hline
\multicolumn{1}{c}{
  \multirow{2}{*}{\textbf{Bug}}} &
  \multicolumn{1}{c}{\multirow{2}{*}{\textbf{\fuzzchick}}} &
  \multicolumn{1}{c}{\multirow{2}{*}{\textbf{\mutagen}}} &
  \multicolumn{2}{c}{\textbf{VD.A}} \\
\cline{4-5}
\multicolumn{1}{c}{} &
\multicolumn{1}{c}{} &
\multicolumn{1}{c}{} &
\multicolumn{1}{c}{\textbf{Value}} &
\multicolumn{1}{c}{\textbf{Estimate}} \\
\hline
1  & 13974.9 & 3632.7 & 0.89 & {\color{red} large} \\ \hline
2  & 23962.9 & 7122.3 & 0.90 & {\color{red} large} \\ \hline
3  & 19678.5 & 4633.9 & 0.89 & {\color{red} large} \\ \hline
4  & 20398.4 & 16831.2 & 0.72 & {\color{red} medium} \\ \hline
5  & 17348.1 & 2326.6 & 0.94 & {\color{red} large} \\ \hline
6  & 10727.1 & 2312.9 & 0.92 & {\color{red} large} \\ \hline
7  & 5070.7 & 332.9 & 0.91 & {\color{red} large} \\ \hline
8  & 5596.4 & 298.7 & 0.90 & {\color{red} large} \\ \hline
9  & 16402.4 & 26342.5 & 0.51 & negligible \\ \hline
10 & 11553.0 & 25044.5 & 0.55 & negligible \\ \hline
11 & 11304.3 & 2536.8 & 0.82 & {\color{red} large} \\ \hline
12 & 18507.3 & 14482.7 & 0.65 & {\color{red} small} \\ \hline
13 & 17961.5 & 13454.9 & 0.65 & {\color{red} small} \\ \hline
14 & 10621.9 & 3928.3 & 0.78 & {\color{red} large} \\ \hline
15 & 23866.3 & 43678.2 & 0.23 & {\color{blue} large} \\ \hline
16 & 25321.7  & 27602.0 & 0.54 & negligible \\ \hline
17 & 28515.6 & 52344.9 & 0.47 & negligible \\ \hline
18 & 24218.6 & 123401.3 & 0.14 & {\color{blue} large} \\ \hline
19 & 18638.9 & 30682.3 & 0.45 & negligible \\ \hline
20 & 18883.1 & 15841.9 & 0.60 & {\color{red} small} \\ \hline
& & \textbf{Mean} & 0.67 & {\color{red} medium} \\
\hline
\end{tabular}
\captionof{table}{\label{table:ifc:tests} Mean tests until first failure and effect size
  comparison between \fuzzchick and \mutagen when considering only succesfull
  runs.
  %
  Estimates in {\color{blue} blue} and {\color{red} red} indicate results
  favouring \fuzzchick and \mutagen, respectively. }
\vspace{-15pt}
\end{wrapfigure}

@@

Table \ref{table:ifc:tests} shows a comparison between the mean number of tests
required by both tools to find the first failure when we only consider
successful runs.
%
Additionally, we compute the non-parametric Vargha-Delaney A measure \cite{vda,
  arcuri}, that encodes the probability of \mutagen to yield better results than
\fuzzchick.
%
As it can be observed, using an exhaustive mutation approach can be not only
more reliable, but also faster, where \mutagen is likely to find bugs faster
that \fuzzchick in 14 of the 20 bugs injected into the abstract machine.

@@

\begin{figure}
  \centering
  \includegraphics[width=\linewidth]{tikz/wasm-mttf.tikz}
  \includegraphics[width=\linewidth]{tikz/wasm-fr.tikz}\vspace{-5pt}
  \includegraphics[width=\linewidth]{tikz/wasm-legend.tikz}
  \caption{\label{fig:results:wasm} Comparison between \quickcheck and \mutagen
    across 15 different bugs for the WebAssembly case study. }
\end{figure}

@@

We consider all these generational approaches to be orthogonal to the ideas
behind \mutagen.
%
In principle, our tool is tailored to improve the performance of poor
automatically derived generators.
%
However, more specialized generators can by used directly by \mutagen, and
combining them with well-typed mutations using a hybrid technique is an
appealing idea that we aim to address in the future.

@@

% SmallCheck and LazySmallCheck
\emph{SmallCheck} \cite{smallcheck} is a Haskell tool that also follows this
idea.
%
It progressively executes the testing properties against all possible inputs
values up to a certain size bound.
%
Later, \emph{LazySmallCheck} is a variant of \emph{SmallCheck} that uses lazy
evaluation to automatically prune the search space by detecting which
subexpressions of the generated values are not evaluated by the testing
properties, and hence do not need to be populated exhaustively with concrete
values.

% Korat
On the object-oriented realm, \emph{Korat} \cite{korat} is a Java tool that uses
method specification predicates to automatically generate all (nonisomorphic)
test cases up to a given small size.
%
Similar to \emph{LazySmallCheck}, \emph{Korat} uses method preconditions to
prune the search space.
%
However, in \emph{Korat}, pruning is done by analyzing the execution trace of
the precondition predicate instead of by keeping track of unused subexpressions.

Being exhaustive, these approaches can only be used to enumerate and test a
relatively little space of small inputs.
%
In this work we use exhaustiveness as a way to reliably enforce that all
possible mutants of an interesting seed are executed.
%
In contrast to fully-exhaustive testing tools, \mutagen relies initially on
randomness to find initial interesting mutable candidates.
%
Later, enumerating mutations for a given interesting candidate is much cheaper
than enumerating \emph{all posible values}.
%
Hence, our tool can be used to produce much larger test cases before their size
becomes computationally too restrictive.

@@

We presented \mutagen, a coverage-guided, property-based testing framework
written in Haskell.
%
Inspired by \fuzzchick, our tool uses coverage information to guide
automatically derived mutators producing high-level, well-typed mutations.
%
However, instead of relying heavily on randomness and power schedules to find
bugs, our tool uses an exhaustive mutation approach that generates every
possible mutant for each interesting input candidate, and schedules it to be
tested exactly once.
%
This is in turn inspired by exhaustive bounded testing tools that focus on
testing every possible input value of the system under test --- a more generic
technique of limited applicability.

Our experimental results indicate that \mutagen outperforms the simpler approach
taken by \fuzzchick in terms of failure rate.
%
Moreover, we show how our tool can be applied in a real-world testing scenario,
where it quickly discovers several planted and existent previously unknown bugs.

>>>>

%%
%% The "title" command has an optional parameter,
%% allowing the author to define a "short title" to be used in page headers.
\title[\mutagen]{\mutagen: Coverage-Guided, Property-Based Testing using
  Exhaustive Type-Preserving Mutations}

@@

%% PBT and generation of interesting data for free
%
Random Property-Based Testing (RPBT) is a popular technique for finding bugs
using executable specifications (properties) encoding the expected behavior of
our system.
%
Exemplified by frameworks Haskell's \quickcheck \cite{quickcheck} and a pletora
of ports of this tool on other programming languages \cite{qcports}, the most
common approach is to instantiate the testing properties using independent,
randomly generated inputs.

Users of \quickcheck (and RPBT in general) are well aware that one of the
biggest challenges while using these frameworks is to provide suitable random
data generators that are used to instantiate the testing properties.
%
In some extreme cases, writing highly-tuned random generators capable of
exercising every part of a complex system in a reasonable basis can take several
thouthands person-hours of trial and error \cite{fuzzchick}.

To alleviate this, there exist several techniques to automatically synthesize
random data generators by reyfing the static information present in the
codebase, e.g., data type definitions, application public interfaces (APIs),
etc.
%
These techniques, however, are unable to synthesize random generators capable of
producing data satisfying complex invariants not easily derivable from the
codebase, being randomly generated programs are common example of this problem.
%
Moreover, automatically synthesized random generators offer particularly poor
results when the testing properties are constrained by sparse preconditions.
%
In this scenario, most of the generated tests are simply discarded before they
can be used to test the main components of our system.

%% FuzzChick: PBT + code instrumentation + type-preserving mutations
\fuzzchick \cite{lampropoulos2019coverage} is a property-based testing framework
for the Coq programming language that borrows ideas from the fuzzing community
to generate highly structured values while using automatically derived
generators.
%
Notably, this tool is implemented as an extension of \quickchick, Coq's own
reimplementation of \quickcheck.

Instead of continuously generating random invalid test cases from scratch,
\fuzzchick keeps a queue of interesting previously executed test cases that can
be mutated using type-preserving transformations in order to produce new ones.
%
Intuitively, mutating an interesting test case in a small way (at the data
constructor level) has a much higher probability of producing a new interesting
test case than generating a new one from scratch using a na\"ive generator.
%
\fuzzchick is likely to preserve the semantic structure of the mutated data, as
mutations are applied directly at the data type level -- the random AST in case
of generating code.

This tool relies on execution traces to distinguish which test cases resulted
interesting and are therefore worth mutating.
%
Mutated test cases are considered interesting for mutation only when they
produce new execution traces -- any other test case is simply thrown away.

Of course, this hybrid technique is not meant to replace smart manually-written
generators.
%
In turn, it provides an acceptable solution that can be used in the early stages
of development, and while a better test suite using manually-written generators
is still under construction.

%% Problems of FuzzChick: mutations are superficial, long queues, power
%% scheduler
%
In this work, we take a close look at \fuzzchick for opportunities to improve.
%
Notably, we establish several aspects of its implementation that can be revised
--- some of these already being identified by the authors of \fuzzchick on their
original work.
%
In particular:
%
\begin{inparaenum}
\item \emph{the type-preserving mutations produced by this tool are superficial
  and rely heavily on randomness to produce diverse mutants},
\item \emph{using normal queues can cause large delays when interesting values
  are enqueued frequently}, and
\item \emph{using a power schedule to assing a certain mutation energy to each
  interesting test case requires fine tuning and can be hard to generalize to
  work well under different testing scenarios.}
\end{inparaenum}

Moreover, we replicated the IFC stack machine case study bundled with \fuzzchick
and observed a noticeable limitation:
%
\emph{after repeating each experiment 30 times, \fuzzchick was only able to find
  5 out of the 20 injected bugs with a 100\% efficacy, the hardest one being
  found only around 13\% of the time after an hour of testing.}
%
Despite this limitation, \citeauthor{lampropoulos2019coverage} show that
\fuzzchick is far better than using normal random testing coupled with na\"ive
random generators by comparing the mean-time-to-failure (MTTF) against
\emph{QuickChick}.
%
However, we believe that a proper evaluation ought to use the failure rate as an
important metric when comparing RPBT tools.

%% Our solution
In this work we introduce \mutagen, a RPBT tool implemented in Haskell that
build upon the basic idea behind \fuzzchick, i.e., code instrumentation and
type-preserving mutations.

%% Exhaustive mutations
Unlike \fuzzchick, mutations in \mutagen are applied \emph{exhaustively}.
%
This is, given an interesting, previously executed test case, our tool
precomputes and schedules every possible type-presserving mutation that can be
applied to it.
%
This approach has two inherent advantages when compared to \fuzzchick.
%
On one hand, mutations do not rely on randomness to be generated.
%
Instead, every subexpression of the input test case is mutated on the same
basis, which guarantees that no interesting mutation is left behind due to
randomness.
%
On the other hand, scheduling mutations exaustively eliminates the need of using
a power schedule to assign a given mutation energy to each input test case.

Notably, \emph{scheduling mutations exhaustively within an input test case does
  not mean that mutations are exhaustive by themselves.}
%
Our tool distinguishes two kinds of mutations, those that can be computed
purely, yielding a single mutated value; and those that can be obtained
non-deterministically.
%
Pure mutations encode transformations that can be done by swapping data
constructor variants or rearranging subexpressions, whereas non-deterministic
(random) mutations are useful to represent transformations over large
enumeration types, e.g., numbers, characters, etc.
%
In this setting, random mutations let us selectively escape the exhaustiveness
of our tool by randomly sampling a small random generator a reduced number of
times.
%
This way, our tool avoids, for instance, mutating an given integer into every
other integer in the 32 bits range.

%% Heuristics to improve testing loop
Additionally, the testing loop of our tool includes implements two heuristics
that help finding bugs faster.
%
In first place, \mutagen uses first-in first-out (FIFO) scheduling with priority
to enqueue interesting test cases for mutation.
%
This scheduling algorithm is indexed by the novelty of the test case with
respect to the ones already seen.
%
In this light, test cases that discover new parts of the code earlier during
execution are given a higher priority while picking the next mutated test case.
%
Additionally, our scheduling lets the testing loop jump back and forth between
scheduled test cases as soon as a new more interesting one is enqueued for mutation,
eliminating potential delays when the mutation queues grow large.

The second heuristic is used to tune the number of times our tool samples random
mutants, i.e., those corresponding to large enumeration types as described above.
%
For this, we keep track of how often we generate interesting test cases.
%
If this frequency suddenly stalls, \mutagen automatically resets the testing
loop increasing the amount of times we sample such mutants.
%
This automatically finds a suitable value for this external on the fly.

%% Validation
To validate our ideas, we use two main case studies (described in detail in
Section \ref{sec:casestudies}): the IFC stack machine used by
\citeauthor{lampropoulos2019coverage} on their original work, and a WebAssembly
engine written in Haskell.
%
In both case studies, we additionally compare the effect of the heuristics
implemented on top of the base testing loop of \mutagen.
%
In the IFC stack machine case study, we compare \fuzzchick against different
variants of our tool, where our results (Section \ref{sec:evaluation}) indicate
that \mutagen outperforms \fuzzchick both in terms of time to failure and
failure rate.
%
On the other hand, our WebAssembly case study shows the performance of our tool
in a more realistic scenario.
%
There, \mutagen is capable of reliaby finding 15 planted bugs in the validator
and interpreter, as well as 3 previously unknown bugs that flew under the radar
of the existing unit test suite of this engine.
%
All in all, our evaluation encompass ?? hours of CPU time and suggest that
testing mutants exhaustively is an appealing technique for finding bugs fast
without sacrificing efficacy.

%% Last two sections

Finally, we discuss related work on Section \ref{sec:related} and conclude
on Section \ref{sec:conclusions}.

@@

In this section we briefly introduce the concept of Property-Based Testing along
with \quickcheck, one of the most popular tools of this sort and often used as
the baseline when comparing PBT algoritms.
%
Moreover, we describe in detail the ideas and limitations behind \fuzzchick,
which served as the foundation while designing \mutagen.

@@

To alleviate the problem of testing properties with sparse preconditions while
using simple automatically derived random generators, \fuzzchick introduces
\emph{coverage-guided property-based testing} (CGPT) by enhancing the testing
process with two key characteristics:
%
\begin{inparaenum}
\item \emph{target code instrumentation}, to capture execution information from
  each test case; and
\item \emph{high-level, type-preserving mutations}, to produce syntactically
  valid test cases by altering existing ones at the datatype level.
\end{inparaenum}

@@

The next subsections describe \fuzzchick's testing loop and type-preserving
mutations in detail.

@@

\subsubsection{Type-preserving mutations}

@@

However, the authors acknowledge that certain parts of its implementation have
room for improvement, especially when it comes to the mutators design.
%
In tandem with the lack of efficacy observed when replicating the evaluation of
the IFC stack machine case study, these observations led us to consider three
main aspects of \fuzzchick that can be improved upon --- and that constitute the
main goal of this work.
%
%% As mentioned earlier, when we recreated the evaluation of the IFC stack machine
%% case study (described in detail in Section \ref{sec:casestudies}), we found that
%% after 30 runs (as opposed to the 5 runs used original by
%% \citeauthor{lampropoulos2019coverage}), \emph{\fuzzchick was only able to find 5
%%   out of the 20 injected bugs with a failure rate of 1}, the hardest one being
%% found only around 13\% of the time after an hour of testing.
%% %
%% These results are presented in detail in Section \ref{sec:evaluation}.
%
%% At the light of these observations, we identified three main aspects of
%% \fuzzchick that can be improved upon --- and that constitute the main goal of
%% this work.
%
In no particular order:
%
\begin{itemize}
\item \emph{Mutators distribution:} if we inspect the mutator defined in Fig.
  \ref{fig:fuzzchick:mutator}, there are two compromises that the authors of
  \fuzzchick adopted for the sake of simplicity.
  %
  On one hand, deep recursive mutations are very unlikely, since their
  probability decreases multiplicatively with each recursive call.
  %
  For instance, mutating a subexpression that lies on the third level of a
  \texttt{Tree} happens with a probability smaller than $(1/6)^3 = \sim 0.0046$,
  and this only worsens as the type of the mutated value becomes more complex.
  %
  Hence, \fuzzchick mutators can only be effectively used to transform to
  shallow data structures, potentially excluding interesting applications that
  might require producing deeper valid values, e.g., programming languages,
  network protocols interactions, etc.
  %
  Ideally, mutations should be able to happen on every subexpression of the
  input seed in a reasonable basis.

@@

\item \emph{Power schedule:} It is not clear how the power schedule used to
  assign energy to each mutable test case in \fuzzchick works in the context of
  high-level type-preserving mutations.
  %
  If it assigns too much energy to certain not-so-interesting seeds, some bugs
  might not be discovered in a timely basis.
  %
  Conversely, assigning too little energy to interesting test cases might cause
  that some bugs cannot be discovered at all unless the right mutation happens
  within the small available energy window --- randomly generating the same test
  case later on does not help, as it becomes uninteresting based on historic
  trace information.

@@

\begin{wrapfigure}{r}{0.5\textwidth}
\centering
\vspace{-10pt}
\footnotesize
\begin{tabular}{ccccc}
\hline
\multicolumn{1}{c}{
  \multirow{2}{*}{Bug}} &
  \multicolumn{1}{c}{\multirow{2}{*}{\fuzzchick}} &
  \multicolumn{1}{c}{\multirow{2}{*}{\mutagen}} &
  \multicolumn{2}{c}{$A_{12}$ measure} \\
\cline{4-5}
\multicolumn{1}{c}{} &
\multicolumn{1}{c}{} &
\multicolumn{1}{c}{} &
\multicolumn{1}{c}{Value} &
\multicolumn{1}{c}{Estimate} \\
\hline
1  & 13974.9 & 3632.7 & 0.89 & {\color{red} large} \\ \hline
2  & 23962.9 & 7122.3 & 0.90 & {\color{red} large} \\ \hline
3  & 19678.5 & 4633.9 & 0.89 & {\color{red} large} \\ \hline
4  & 20398.4 & 16831.2 & 0.72 & {\color{red} medium} \\ \hline
5  & 17348.1 & 2326.6 & 0.94 & {\color{red} large} \\ \hline
6  & 10727.1 & 2312.9 & 0.92 & {\color{red} large} \\ \hline
7  & 5070.7 & 332.9 & 0.91 & {\color{red} large} \\ \hline
8  & 5596.4 & 298.7 & 0.90 & {\color{red} large} \\ \hline
9  & 16402.4 & 26342.5 & 0.51 & negligible \\ \hline
10 & 11553.0 & 25044.5 & 0.55 & negligible \\ \hline
11 & 11304.3 & 2536.8 & 0.82 & {\color{red} large} \\ \hline
12 & 18507.3 & 14482.7 & 0.65 & {\color{red} small} \\ \hline
13 & 17961.5 & 13454.9 & 0.65 & {\color{red} small} \\ \hline
14 & 10621.9 & 3928.3 & 0.78 & {\color{red} large} \\ \hline
15 & 23866.3 & 43678.2 & 0.23 & {\color{blue} large} \\ \hline
16 & 25321.7  & 27602.0 & 0.54 & negligible \\ \hline
17 & 28515.6 & 52344.9 & 0.47 & negligible \\ \hline
18 & 24218.6 & 123401.3 & 0.14 & {\color{blue} large} \\ \hline
19 & 18638.9 & 30682.3 & 0.45 & negligible \\ \hline
20 & 18883.1 & 15841.9 & 0.60 & {\color{red} small} \\ \hline
& & Mean & 0.67 & {\color{red} medium} \\
\hline
\end{tabular}
\captionof{table}{\label{table:ifc:tests} Mean tests until first failure and effect size
  comparison between \fuzzchick and \mutagen when considering only succesfull
  runs.
  %
  Estimates in {\color{blue} blue} and {\color{red} red} indicate results
  favouring \fuzzchick and \mutagen, respectively. }
\vspace{-15pt}
\end{wrapfigure}

@@

Table \ref{table:ifc:tests} shows a comparison between the mean number of tests
required by both tools to find the first failure when we only consider
successful runs.
%
Additionally, as adviced by \cite{arcuri} when comparing random testing tools,
we computed the non-parametric Vargha-Delaney $A\_{12}$ measure \cite{vda}.
%
Intuitively, this measure encodes the probability of \mutagen to yield better
results than \fuzzchick.
%
As it can be observed, using an exhaustive mutation approach can be not only
more reliable, but also faster, where \mutagen is likely to find bugs faster
that \fuzzchick in 14 of the 20 bugs injected into the abstract machine.

@@

\begin{figure}
  \centering
  \includegraphics[width=\linewidth]{tikz/wasm-mttf.tikz}
  \includegraphics[width=\linewidth]{tikz/wasm-fr.tikz}\vspace{-5pt}
  \includegraphics[width=\linewidth]{tikz/wasm-legend.tikz}
  \caption{\label{fig:results:wasm} Comparison between \quickcheck and \mutagen
    across 15 different bugs injected into \textit{haskell-wasm}. }
\end{figure}

@@

We consider all these generational approaches to be orthogonal to the ideas
behind \mutagen.
%
In principle, our tool is tailored to improve the performance of poor
automatically derived generators.
%
However, more specialized generators can by used directly by \mutagen, and
combining them with type-preserving mutations using a hybrid technique is an
appealing idea that we aim to address in the future.

@@

% SmallCheck and LazySmallCheck
\emph{SmallCheck} \cite{smallcheck} is a Haskell tool that also follows this
idea.
%
It progressively executes the testing properties against all possible inputs
values up to a certain size bound.

% Korat
On the object-oriented realm, \emph{Korat} \cite{korat} is a Java tool that uses
method specification predicates to automatically generate all (nonisomorphic)
test cases up to a given small size.

Being exhaustive, these approaches rely on prunning mechanisms to avoid
populating unevaluated subexpresions exhaustively before the computational cost
becomes too restrictive.
%
\emph{LazySmallCheck} is a variant of \emph{SmallCheck} that uses lazy
evaluation to automatically prune the search space by detecting unevaluated
subexpressions using lazy evaluation.
%
In the case of \emph{Korat}, pruning is done by instrumenting method
precondition predicates and analyzing which parts of the execution trace
correspond to each evaluated subexpression.

In this work we use exhaustiveness as a way to reliably enforce that all
possible mutants of an interesting seed are executed.
%
In contrast to fully-exhaustive testing tools, \mutagen initially relies on
randomness to find initial interesting mutable candidates.
%
In terms of pruning, it is possible to instruct \mutagen to detect unused
subexpressions (using a technique similar to that of \emph{LazySmallCheck}) to
avoid scheduling mutations over their corresponding positions.
%
This could, in principle, improve the overall performance when testing
properties that tend to short-circuit, leaving parts of the input unevaluated.
%
In our case studies, however, we observed that their preconditions tend to be
quite strict when mutating from passed test cases, fully evaluating their input
before executing the postcondition.
%
Thus, we do not expect to see considerable improvements by applying this idea in
this particular scenario.
%
Gathering empirical evidence about using prunning via lazy evaluation in
\mutagen is an effort that we keep as future work.

@@

We presented \mutagen, a coverage-guided, property-based testing framework
written in Haskell.
%
Inspired by \fuzzchick, our tool uses coverage information to guide
automatically derived mutators producing high-level, type-preserving mutations.
%
However, instead of relying heavily on randomness and power schedules to find
bugs, our tool uses an exhaustive mutation approach that generates every
possible mutant for each interesting input candidate, and schedules it to be
tested exactly once.
%
This is in turn inspired by exhaustive bounded testing tools that focus on
testing every possible input value of the system under test --- a more generic
technique of limited applicability.

Our experimental results indicate that \mutagen outperforms the simpler approach
taken by \fuzzchick in terms of both failure rate and tests until first failure.
%
Moreover, we show how our tool can be applied in a real-world testing scenario,
where it quickly discovers several planted and existent previously unknown bugs.
